### 2023-03-24 paper/tables/webdata.tex
initialize paper

\begin{table*}
    \centering
    \begin{tabular}{lcccc}
        \hline
        \textbf{Treebank} & \textbf{Language} & \textbf{N. sentences} & \textbf{Abbreviation}\\
        \hline
        UD English-ESL/TLE & English & 4124 & en\_esl \\
        \hline
        UD English-GUM & English & 6911 & en\_gum \\
        \hline
        UD English-EWT & English & 12544 & en\_ewt \\
        \hline
        UD German-GSD & German & 13814 & de\_gsd \\
        \hline
        UD German-HDT & German & 75617 & de\_hdt \\
        \hline
        UD Chinese-GSD & Chinese & 3997 & cmn\_gsd \\
        \hline
        UD Chinese-PUD & Chinese & 1000 & cmn\_pud \\
        \hline
        UD Hungarian-Szeged & Hungarian & 455, 455 & hun\_szeged\_1, hun\_szeged\_2 \\
        \hline
    \end{tabular}
    \caption{UD web treebanks and sources}
    \label{tab:webdata}
\end{table*}