### 2023-03-24 paper/tables/metrics.tex

\begin{table*}[ht]
    \centering
    \begin{tabular}{p{100px}p{150px}cc}
        \hline
        \textbf{Metric} & \textbf{Description} & \textbf{Abbreviation} & \textbf{Level} \\
        \hline
        Token Count & Token length of sentence & \emph{tc} & sentence \\
        \hline
        Type Token Ratio & Simple type token ratio, without preprocessing & \emph{ttr} & sentence \\
        \hline
        Parse Tree Depth & Depth of the parse tree obtained from the sentence & \emph{ptd} & sentence \\
        \hline
        Length of longest dependency link & Surface distance in tokens between the most distant related elements in a sentence & \emph{lldl} & sentence \\
        \hline
        Noun to verb ratio & Number of nouns divided by the number of verbs in a sentence & \emph{n2v} & sentence \\
        \hline
        Number of clauses per sentence & Number of clauses per sentence - clauses are identified by the presence of verbs & \emph{cxc} & sentence \\
        \hline
        Number of xcomp relations per sentence & Number of xcomp relations per sentence & \emph{xcomp} & sentence \\
        \hline
        Number of ccomp relations per sentence & Number of ccomp relations per sentence & \emph{ccomp} & sentence \\
        \hline
        Ratio of verbs with explicit subject & Number of verbs in a sentence with explicit \emph{nsubj} or \emph{csubj} relations, divided by the number of verbs & \emph{ves} & sentence \\
        \hline
        Word forms per lemma & Number of UD FORM variants for every LEMMA & \emph{wfpl} & treebank \\
        \hline
        POS sequence variability & Variability of sequences of three POS tags & \emph{vsp} & treebank \\
        \hline
    \end{tabular}
    \caption{Complexity metrics}
    \label{tab:metrics}
\end{table*}