### 2023-03-25 paper/tables/results_ours.tex
update paper

\begin{table*}
\centering
\begin{tabularx}{\textwidth}{Xcccccc}
\hline
\textbf{Treebank} & \textbf{tc} & \textbf{ttr} & \textbf{ptd} & \textbf{lldl} & \textbf{n2v} & \textbf{cxc} \\
\hline
en & \textbf{14.6} & 0.9264 & 3.58 & \textbf{12.44} & 1.6202 & 1.54 \\
\hline
deu & 13.38 & 0.9311 & 3.38 & 11.44 & 1.6287 & 1.18 \\
\hline
hun & 10.8 & \textbf{0.9604} & 3.2 & 6.68 & \textbf{1.711} & 1.3 \\
\hline
cmn & 14.52 & 0.9117 & \textbf{3.66} & 12.1 & 1.2491 & \textbf{1.96} \\
\hline
\end{tabularx}

\vspace{0.5cm}

\begin{tabularx}{\textwidth}{Xcccccc}
\hline
\textbf{Treebank} & \textbf{xcomp} & \textbf{ccomp} & \textbf{ves} & \textbf{ves\_t} & \textbf{wfpl} & \textbf{vps} \\
\hline
en & 0.14 & 0.16 & 0.6117 & 0.6548 & 1.2075 & 311.0 \\
\hline
deu & 0.16 & 0.14 & \textbf{0.6267} & \textbf{0.75} & 1.1977 & 305.0 \\
\hline
hun & 0.1 & \textbf{0.24} & 0.3417 & 0.378 & \textbf{1.2524} & 236.0 \\
\hline
cmn & \textbf{0.5} & 0.22 & 0.4892 & 0.4745 & 1.0051 & \textbf{330.0} \\
\hline
\end{tabularx}
\caption{Results on in-house parallel treebank. Top values are highlighted for each metric.}
\label{tab:results-ours}
\end{table*}